\begin{acknowledgements}

I would be remiss not to begin by expressing my deep gratitude to Oxford for the vast natural beauty and awe-inspiring architecture it has blessed me with in the last few months. My time here has changed me in ways I have not yet begun to process, akin to the grandeur of altered experiences embedded in the pages of this thesis. To Jan, for the trust you placed in me. Overwhelmed, I stepped into your office;  with just a few words you introduced me to scientific passions I had no idea were within me. To Morten, for welcoming me with open arms into the Eudaimonia community, and for teaching me scientific rigor, curiosity, criticality, and wisdom. To my family, for the excitement they have for my journey through this academic life. Our lives are so different, and yet you have always shown me unconditional support and understanding.

To the wider Eudaimonia community for their generous and gracious support, advice, and friendship. Thank you to Leonardo for solving every silly roadblock I ran into in the darkness of the back office. To Kat, Elvira, Gustavo, Dimitri, Pedro, Fernando, Shamil, Robin, and Joana for the kindness, questions, and critique which allowed me to repair the cracks in my work, and my thinking about the brain and mind.

Thank you to the friends I have made since my time in Oxford, whose limitless optimism, curiosity, and love have changed the way I think about friendship, science, and life, and kept me grounded as my mind tripped and fell over the complexities of entropy and hierarchy. Thank you to Michael, Kenneth, Fil, Clayton, Will, Elin, Sebastian, Victoria, and Marilena, whose presence has given me more than can be expressed in words, or at least the limited space I have here.

Last but not least, to Maastricht and the psychopharmacology team. To Pablo for holding my hand with patience as I took my first steps in the blinding, expansive world of neuroimaging. A summer I very well could have spent traveling was spent behind my laptop, bashing keys with frustration as FSL threw errors for the thousandth time. I would not change that for anything. Thank you to Petr, who saw the light in me and challenged me to let it out. And to Baptiste, Mila, and Aedan for every moment shared, whether a long night in the library or in Cafe Zondag.
\end{acknowledgements}